---
title: "The Factors Affect CO2 Emission in 5-ASEAN Countries"
author: "Auggie Imuan"
date: "2021-3-31"
categories: [economics, analysis]
---

## 1. Introduction

Global warming is driving significant changes in the climate system. The primary cause of global warming is the accumulation of greenhouse gases (GHG). A growing amount of GHG emissions can be attributed to human activities, such as industrial expansion, transportation, and other economic activities. Among these, carbon dioxide (CO2) emissions are the largest contributors to total GHG emissions (Intergovernmental Panel on Climate Change, 2014). CO2 constitutes more than 72% of global GHG emissions (Yunfeng and Laike, 2010). Furthermore, the CO2 emission share from developing countries was nearly 50 per cent of the world's total CO2 emissions in 2003 (Martinez-Zarzoso and Maruotti, 2011).

A significant proportion of developing countries is represented within ASEAN, a region characterized by rapid economic development, rising income levels, and other socio-economic drivers. Increased domestic and foreign investments in ASEAN nations have led to a continuous rise in CO2 emissions, contributing to environmental degradation.

Although ASEAN's CO2 emissions are lower than those of China and the USA, ASEAN is predicted to face severe consequences of global warming, including increasing sea levels (Nordin et al., 2015), which endangered coastal areas and small islands, and the severity of extreme weather events. The population in ASEAN are affected by CO2 emissions because the region is populated along coastlines and heavily reliant on agriculture for livelihoods. Rising temperatures caused by increased CO2 emissions directly affect natural resources and forestry, leading to declining rainfall, heatwaves, droughts, floods, and tropical cyclones (Environment Division of the ASEAN Secretariat, n.d.). 

The increase in CO2 emissions in ASEAN has tended to increase occasionally. In 2018, the 5 countries in ASEAN that had the most CO2 emissions were Indonesia, Thailand, Vietnam, Malaysia, and the Philippines, respectively.

```{r}
#| echo: false
#| warning: false
#| include: false

library(tidyverse)
library(ggplot2)
library(plotly)

co2 <- read_csv("data/co2.csv")
energy <- read_csv("data/energy.csv")
gdp <- read_csv("data/gdp.csv")
industry <- read_csv("data/industry.csv")
population <- read_csv("data/population.csv")
urban <- read_csv("data/urban.csv")
ASEAN_co2 <- read_csv("data/co2_ASEAN.csv")




clean_dataframe <- function(data) {
  # Remove the "Series Code" column 
  if ("Series Code" %in% colnames(data)) {
    data <- data[, !colnames(data) %in% "Series Code"]
  }
  
  # Rename "Series Name" column to "Variable"
  colnames(data) <- ifelse(colnames(data) == "Series Name", "Variable", colnames(data))
  
  # Clean column names to remove brackets and whitespace
  colnames(data) <- gsub("\\[.*\\]", "", colnames(data)) # Remove anything in square brackets
  colnames(data) <- trimws(colnames(data)) # Remove leading and trailing whitespace
  
  # Select only rows 1 to 5
  data <- data[1:5, ]
  
  # Ensure all year columns are numeric
  numeric_columns <- colnames(data)[!colnames(data) %in% c("Country Name", "Country Code", "Variable")]
  data[numeric_columns] <- lapply(data[numeric_columns], function(x) as.numeric(as.character(x)))
  
  # Pivot the year columns into a long format
  data <- pivot_longer(
    data,
    cols = -c(`Country Name`, `Country Code`, `Variable`), # All columns except these
    names_to = "Year", # Name for the new "Year" column
    values_to = "Value" # Name for the new "Value" column
  )
  
  # Return data frame
  return(data)
}

clean_dataASEAN <- function(data) {
  # Remove the "Series Code" column 
  if ("Series Code" %in% colnames(data)) {
    data <- data[, !colnames(data) %in% "Series Code"]
  }
  
  # Rename "Series Name" column to "Variable"
  colnames(data) <- ifelse(colnames(data) == "Series Name", "Variable", colnames(data))
  
  # Clean column names to remove brackets and whitespace
  colnames(data) <- gsub("\\[.*\\]", "", colnames(data)) # Remove anything in square brackets
  colnames(data) <- trimws(colnames(data)) # Remove leading and trailing whitespace
  
  # Select only rows 1 to 10
  data <- data[1:10, ]
  
  # Ensure all year columns are numeric
  numeric_columns <- colnames(data)[!colnames(data) %in% c("Country Name", "Country Code", "Variable")]
  data[numeric_columns] <- lapply(data[numeric_columns], function(x) as.numeric(as.character(x)))
  
  # Pivot the year columns into a long format
  data <- pivot_longer(
    data,
    cols = -c(`Country Name`, `Country Code`, `Variable`), # All columns except these
    names_to = "Year", # Name for the new "Year" column
    values_to = "Value" # Name for the new "Value" column
  )
  
  # Return data frame
  return(data)
}


my_var <- c("co2", "energy", "gdp", "industry", "population", "urban")

# Loop through each data frame name
for (df_name in my_var) {
  # Get the data frame by its name
  df <- get(df_name)
  
  # Apply the cleaning function
  cleaned_df <- clean_dataframe(df)
  
  # Assign the cleaned data frame back to the original variable
  assign(df_name, cleaned_df)
}

ASEAN_co2 <- clean_dataASEAN(ASEAN_co2)
```

```{r}
#| echo: false
#| warning: false
ASEAN_co2 <- ASEAN_co2 %>% filter(!is.na(Value)) %>%
  mutate(Year = as.numeric(Year))

ggplot(ASEAN_co2, aes(x = Year, y = Value, color = `Country Name`, group = `Country Name`)) +
  geom_line(size = 1) +  # Slightly thicker lines for better visibility
  geom_point(size = 0.5) +   # Larger points for emphasis
  labs(
    title = "CO2 Emissions across ASEAN countries",
    subtitle = "CO2 Emissions (1990-2022)",
    x = "Year",
    y = "Emissions (in Metric Tons)",
    color = "Country"
  ) +
  scale_x_continuous(breaks = seq(1970, 2022, 5)) + 
  theme_bw() +  # Cleaner theme
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),  # Centered and bold title
    plot.subtitle = element_text(hjust = 0.5, size = 12),  # Centered subtitle
    axis.title = element_text(size = 14),  # Larger axis titles
    axis.text = element_text(size = 12),   # Larger axis text
    axis.text.x = element_text(angle = 45, hjust = 1),  # Rotate x-axis labels
    legend.title = element_text(size = 12),  # Adjust legend title size
    legend.text = element_text(size = 10),   # Adjust legend text size
    panel.grid.major = element_line(color = "gray", linetype = "dotted"),  # Dotted grid lines
    panel.grid.minor = element_blank()  # Remove minor grid lines
  )

```

Figure 1 can be implied that the countries top five CO2 emission in ASEAN have environmental problems. Moreover, this figure implied about what the factors that make countries top five CO2 emission in ASEAN. To achieve sustainable economic development, their countries should be taken certain step to stop or control CO2 emission (Abdul and Baserat, 2018). As Grossman and Krueger were the first to introduce the idea of the environmental Kuznets curve (EKC) hypothesis to show a negative relationship between economics growth and environmental quality (Arifur et al., 2020). And in 1971, Ehrlich and Holdren were the first introduce the idea of the analysis of key human activities that influence environmental change that is environmental impacts (I) are associated with the population (P), affluence (A), and technology (T) in ecological model called “IPAT” (Koçak et al., 2020). From 2 hypotheses, which are appropriate and can most explain factors affecting CO2 emission on the top 5 CO2 emission in ASEAN is IPAT hypothesis because IPAT theory is the effective device for operationally bridging the differing perspectives between social science and biological science (Dietz and Rosa, 1994). While the EKC ignored the impact of outside borders on the environment since the environment cannot be isolated, the IPAT model assumes the environmental impacts increase linearly with technology (Majeed and Tauqir, 2020).

However, this problem has not been directly studied, but studies on determinate of CO2 emission in ASEAN by using energy indicators (Nordin et al., 2015), and studies by Qaiser et al (2020) on relationship between CO2 emission, energy consumption and economic growth in 5-ASEAN (i.e. Indonesia, Thailand, Singapore, Malaysia, and Philippines) by using the EKC hypothesis. Therefore, this research has the objective to study applies IPAT hypothesis to explore factors affecting CO2 emission on top 5 CO2 emission in ASEAN.


## 2. Literature Review
### 2.1 Theory of Environmental Degradation

Nowadays, the most important problem in the world is the environmental problem. The main reason for the environmental degradation problem due to human activities. Therefore, scientists study and create theories, to explain how human activities affect the degradation of the environment. The popular theories, that describe environmental degradation, are the EKC theory and IPAT theory (Carson, 2010).   

The EKC theory, or the environmental Kuznets curve, was created by Simon Smith Kuznets, an economist and statistician. This theory describes how the amount of CO2 emissions will increase until the average income reaches a turning point, thus revitalizing environmental quality. Furthermore, the conventional EKC illustrates an inverted U-shaped relationship between environmental quality and economic growth/development (Grossman and Krueger, 1995).

On the contrary, the IPAT theory or The Environmental Impacts (I) on Population (P), Affluence (A), and Technology (T) created by Paul Ralph Ehrlich, Barry Commoner, and John Paul Holdren who are biologists, ecologists, physics and environmental scientist, respectively (York et al., 2003). The IPAT model postulates that environmental impacts increase linearly with population and affluence (York et al., 2007). Moreover, the purpose of this theory is to illustrate the population is a key driving force. It shows humans are important for driving/creating economic activities, technology, political and economic institutions, and attitudes (Stem et al., 1992). 

In that case, the IPAT theory is better than the EKC theory, which describes reasons that affect environmental degradation. Dietz and Rosa (1997) believe IPAT theory is an effective device for operationally bridging the differing perspectives between social science and biological science. While the EKC ignored the impact of outside borders on the environment since the environment cannot be isolated, the IPAT model assumes the environmental effects increase linearly with technology (Majeed and Tauqir, 2020). Moreover, biological, ecological and other physical and environmental scientists' models do not ignore assumptions and create dependence on actuality, while some social science will generally overlook the models (Dietz and Rosa, 1994). Mill and Waite's (2009) studies did not support EKC's role of affluence and population on the environment.

### 2.2 Factors affects to CO2 emission

As humans and human activities are the main reasons in the theory of environmental degradation. Even though environmental degradation cannot be directly measured, most research studies on environmental degradation will use CO2 emission to measure it (York et al., 2003).

Undoubtedly, the population is a significant reason for CO2 emissions due to the impact of the increasing population leading to the use of more energy and oil consumption even with the most managed technology (Paul and John, 1971). Moreover, Mansoor and Sultana (2018), and Guo and Yuan (2011) investigated the increasing population effect on CO2 emission, which led to environmental degradation. While increasing the population concentrated in urban centres, it began to expand urban areas (Lin et al., 2009). This made the migration of the rural population to urbanization thus using more energy and oil consumption led to environmental degradation (Yeh and Liao, 2017). Some studies such as Yeh and Liao (2017), Le et al (2020), Martinez-Zarzoso and Maruotti (2011), and Shaari et al (2020) questioned the urbanisation effect on carbon emission by using the IPAT/STIRPAT tool to analyze, found the relation of urbanization with carbon emission is positive linear.

On the contrary, Le and Quah (2018) and Husain (2016) questioned the urbanisation effect on carbon emission by using the EKC model to analyse and found the relation of urbanization with carbon emission is a U-shape. In other words, low levels of urbanization cause more CO2 emission while levels of urbanization reach a point that will reduce CO2 emission.

The other cause of CO2 emission is human activities, which can be divided into 2 types followed by IPAT theory including affluence and technology (York et al., 2003) which will mention affluence and technology, respectively. Affluence is the production/consumption of the population and affected environment, the best of indicators are income and industrial. Firstly, income can be measured by power consumption. The income illustrates the power of the consumer to afford the product, which is the cause of CO2 emission. When people can more afford the product, it leads to more production which leads to increasing CO2 emission (Guo and Jiang, 2011). In addition, Yeh and Liao (2017) and Le et al (2020) questioned the gross domestic product effect on carbon emission by using the IPAT/STIRPAT tool to analyse and found the relation of gross domestic product with carbon emission is positive linear. Some studies such as Giedre et al (2014), Le and Quah (2018), Husain (2016), Sulaiman and Abdul-Rahim (2018) and Neumayer (2002) estimated the gross domestic product effect on carbon emission by using the EKC model to analyze, found the relation of gross domestic product with carbon emission is positive as well. 

Although most studies investigate the positive relationship between gross domestic product and carbon emission, Mansoor and Sultana (2018) investigate that GDP per capita has a negative effect on CO2 emission because GDP enables a country to reach the same production level at a low CO2 emission by development of low-carbon technology. Thus the indicator can be measured, production is industrial. Undoubtedly, CO2 emission comes from production in industry. If countries have more industry, which means increasing production, they have to use energy and oil consumption in the production process which leads to more CO2 emissions. Moreover, industrial illustrates international trade, investment, urbanization and economic growth (Le and Quah, 2018; Saboori et al., 2017). Lin et al (2009) estimated an analysis of the environmental impact of China based on the STIRPAT model. This study found the relationship between the level of industry and CO2 emission is positive, in contrast, Paramati et al (2016) examined developed economy countries that have a negative relationship between industry and CO2 emission because developed economy countries have a high level of technology to reduce CO2 emission.

Finally, the last cause of CO2 emission by human activities is technology. Paul and John (1971) believe that in a technological society, each human individual has a negative impact on the environment. Even though, technology is an indicator that difficult measures to value, most researchers use energy consumption to describe the level of technology, for example, if less energy consumption refers to the level of technology is high (Mansoor and Sultana, 2018). Zhang and Chenh (2009) examined how an increase in energy consumption brings about an increased CO2 emission. In addition, the indicators that use rather energy consumption is energy intensity (Koçak et al., 2020). Wolde-Rufael and Idowu (2017) investigated that China and India have a positive relationship between energy consumption and CO2 emission due to low-level technology. Thus it used high energy intensity in contrast low energy intensity implied the high level of technology. Talbi (2017) and Le et al (2020) express energy consumption illustrate technology, moreover; energy consumption measured by energy intensity investigated energy intensity positive effects on CO2 emission in the short run. On the contrary, Koçak et al (2020) and Xu and Ang (2013) found the reduction in energy intensity represents technological development, which is the key driving force for reducing emissions, and found energy intensity has a negative relationship with CO2 emission.    





